---
title: "Coffee Shop Revenue Analysis"
author: "Pranav Jayakumar"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

## Dataset Information
- **Dataset Link**: [https://www.kaggle.com/datasets/himelsarder/coffee-shop-daily-revenue-prediction-dataset]  
- This dataset contains daily revenue data for a coffee shop, including:  
  - Number of customers per day  
  - Average order value  
  - Operating hours  
  - Marketing spend  
  - Foot traffic  
  - Daily revenue  

## Motivation
Understanding what factors drive daily revenue is crucial for optimizing business strategies. This analysis aims to:  
- Identify key revenue drivers (customer count, order value, etc.).  
- Assess how foot traffic correlates with customers.  
- Evaluate marketing effectiveness.  

## Summary Statistics
```{r}
library(dplyr)
library(ggplot2)

df <- read.csv("coffee_shop_revenue.csv")
summary(df)
```

# Visualizations

## Relationship between Two Quantitative Variables
```{r}
ggplot(df, aes(x = Number_of_Customers_Per_Day, y = Daily_Revenue)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Customers vs Revenue",
       x = "Number of Customers Per Day",
       y = "Daily Revenue")
```
**Interpretation:**  
- There is a strong positive relationship between the number of customers and revenue.  
- More customers generally lead to higher daily revenue.  

## Relationship between a Categorical and Quantitative Variable
```{r}
df$Customers_High_Low <- ifelse(df$Number_of_Customers_Per_Day > median(df$Number_of_Customers_Per_Day), "High", "Low")
ggplot(df, aes(x = Daily_Revenue, fill = Customers_High_Low)) +
  geom_density(alpha = 0.5) +
  labs(title = "Revenue Distribution: High vs. Low Customer Count",
       x = "Daily Revenue",
       y = "Density")
```
**Interpretation:**  
- High-customer days tend to have significantly higher revenue compared to low-customer days.  

# Dataset Manipulations

## Using dplyr for Data Transformations
```{r}
df %>%
  group_by(Customers_High_Low) %>%
  summarize(Average_Revenue = mean(Daily_Revenue)) %>%
  ggplot(aes(x = Customers_High_Low, y = Average_Revenue, fill = Customers_High_Low)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Revenue by Customer Group",
       x = "Customer Count Group",
       y = "Average Revenue")
```
**Insight:**  
- The high customer count group consistently generates higher average revenue.  

# Statistical Analyses

## Correlation Analysis
```{r}
library(ggcorrplot)
cor_matrix <- cor(df %>% select_if(is.numeric), use = "complete.obs")
ggcorrplot(cor_matrix, method = "circle", type = "lower", lab = TRUE,
           title = "Correlation Matrix")
```
**Interpretation:**  
- Number of customers strongly correlates with daily revenue.  
- Order value has a moderate correlation with revenue.  
- Surprisingly, foot traffic has little to no effect on the number of customers or daily revenue, suggesting that customers generally do not go to coffee shops unprompted.  

## Linear Regression
```{r}
lm_model <- lm(Daily_Revenue ~ Number_of_Customers_Per_Day, data = df)
summary(lm_model)
```
**Interpretation:**  
- Each additional customer increases revenue by about $5.55.  
- The model predicts an average of $393.86 of daily revenue with no customers (possibly due to memberships, standing orders, etc.).  
- The model explains a moderate portion of the model variation.  

## T-Test
```{r}
t_test_result <- t.test(Daily_Revenue ~ Customers_High_Low, data = df)
t_test_result
```
**Interpretation:**  
- High-customer days result in significantly higher revenue.  
- The difference is statistically significant (p-value < 0.05).  

# Conclusion
- Customer count is the strongest predictor of revenue.  
- Order value also contributes but is less influential.  
- Marketing strategies should focus on increasing customer traffic.  
- Upselling strategies can help boost order value.  

**Future Considerations:**  
- Explore seasonal effects and promotions.  
- Use a multiple regression model to refine predictions.  
